---
title: "EDA-zhou"
output: html_document
---

EDA for the train dataset 




```{r}
#load library
library(ggplot2)
library(dplyr)
library(readr)
library(ggcorrplot)
library(GGally)
library(FSA)
```


```{r}
#load dataset
train <- read.csv("C:\\Users\\leonz\\Desktop\\project\\InsNova_train.csv")

train.df <- train
```



#Overall data structure
```{r}
head(train.df)
```

```{r}
summary(train.df)
```
check NAs
```{r}
sum(is.na(train.df))
```
no NAs in this dataset

#convert all chatracter variables to factor
```{r}
train.df[sapply(train.df, is.character)] <- lapply(train[sapply(train.df, is.character)], 
                                       as.factor)

```

```{r}
summary(train.df)

```


numeric variabbles: 
veh_value 
exposure  
veh_age
dr_age


facttor variables: 
veh_body 
gender
area  

#Numerica variable EDA
##veh_value 
```{r}
hist(train.df$veh_value)
```

```{r}
summary(train.df$veh_value)
```
```{r}
boxplot(train.df$veh_value)
```
```{r}
plot(claim_cost ~ veh_value, data = train.df)
```
##exposure

```{r}
hist(train.df$exposure)
```

```{r}
summary(train.df$exposure)
```
```{r}
boxplot(train.df$exposure)
```
```{r}
plot(claim_cost ~ exposure, data = train.df)
```

##veh_age
```{r}
hist(train.df$veh_age)
```
```{r}
summary(as.factor(train.df$veh_age))

```

```{r}
plot(claim_cost ~ veh_age, data = train.df)
```

```{r}
boxplot(claim_cost ~ as.factor(veh_age), data = train.df)
```
```{r}
Summarize(claim_cost ~ as.factor(veh_age),
          data=train.df)
```



##dr_age
```{r}
hist(train.df$dr_age)
```

```{r}
summary(as.factor(train.df$dr_age))

```
```{r}
boxplot(claim_cost ~ as.factor(dr_age), data = train.df)
```


```{r}
Summarize(claim_cost ~ as.factor(dr_age),
          data=train.df)
```


##claim_cost
```{r}
hist(train.df$claim_cost)
```
number of 0s
```{r}
sum(train.df$claim_cost==0)
```
non 0s
```{r}
sum(train.df$claim_cost!=0)
```

```{r}
cost_sub <- subset(train.df, claim_cost!=0)
```

```{r}
hist(cost_sub$claim_cost)
```
```{r}
summary(cost_sub$claim_cost)
```

##correlational plot between numerical variables

```{r}
num_sub <- train.df[, c("veh_value", "exposure", "veh_age", "dr_age","claim_cost")]

```



```{r}
cor(num_sub)
```
```{r}
corr <- round(cor(num_sub), 0.00001)
ggcorrplot(corr)
```


#Categorical variable 

##veh_body
```{r}
summary(train.df$veh_body)
```
```{r}
Summarize(claim_cost ~ veh_body,
          data=train.df)
```



```{r}
boxplot(claim_cost ~ veh_body,
          data=train.df)
```


##gender 

```{r}
summary(train.df$gender)
```

```{r}
Summarize(claim_cost ~ gender,
          data=train.df)

```

```{r}
boxplot(claim_cost ~ gender,
          data=train.df)
```

##area
```{r}
summary(train.df$area)
```

```{r}
Summarize(claim_cost ~ area,
          data=train.df)

```

```{r}
boxplot(claim_cost ~ area,
          data=train.df)
```