---
title: "Lab1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




Agenda: 
- download and install R on your computer 

- install R packages 

 - get familiar with R 

- load essential packages and datasets for the course 


R installation: https://mirror.las.iastate.edu/CRAN/
RStudio installation: https://www.rstudio.com/products/rstudio/download/

*R basics
```{r}
x <- c(10,3,3,4)
x + 4
```

```{r}
y <- max(x)
y
```

```{r}
x^2
```



```{r}
y <- min(x)
y 
```


Working with Data Objects
```{r}
x <- c(1,3,4,6,5,6)
x
```

```{r}
mean(x) #mean
```


```{r}
sd(x) #standard deviation 
```


```{r}
var(x)
```


Next, create matrix 
```{r}
y <- matrix(x, nrow = 3)
y
```
math operations
```{r}
exp(1)
log(exp(4))
sqrt(20)
```

R cheatsheets:
https://www.rstudio.com/resources/cheatsheets/



Loading Some Essential Packages

Data on Porsche prices
```{r}
Install.packages(Stat2Data)
library(Stat2Data)
Data(PorschePrice)
attach(PorschePrice)
PorschePrice
```
Data on Doctors and hospitals in metropolitan areas 
```{r}
library(Stat2Data)
Data(MetroHealth83)
attach(MetroHealth83)
MetroHealth83
```

Data on   Olympics Long Jump
```{r}
library(Stat2Data)
Data(LongJumpOlympics)
attach(LongJumpOlympics)
LongJumpOlympics
```


Data on Butterfly Ballot
```{r}
Library(Stat2Data)
Data (PalmBeach )
attach(PalmBeach)
PalmBeach
```

 Sparrows Data
```{r}
library(Stat2Data)
Data(Sparrows)
attach(Sparrows)
Sparrows
```

breakfast cereal Data
```{r}
library(Stat2Data)
Data (Cereal)
attach(Cereal)
Sparrows
```

Capacitor voltage Data
```{r}
library(Stat2Data)
Data (Volts)
attach(Volts)
Volts
```

Caterpillars Data
```{r}
library(Stat2Data)
Data (Caterpillars)
attach(Caterpillars)
Caterpillars
```





