---
title: "pubh7450-assign0"
output: html_document
---

```{r}
library(ggplot2)
```

#1.load data
```{r}
chicks <- read.csv('C:/Users/leonz/Desktop/UMN/Fall 2021/PubH7430 - Statistical methods for correlated data/Assignment/Assignment 0/chicks.csv')
head(chicks)
chicks$Chick <- as.factor(chicks$Chick)
chicks$Diet <- as.factor(chicks$Diet)
```






#2. 
```{r}
class(chicks$Chick)
length(unique(chicks$Chick))
```

There are 50 chicks 


```{r}
summary(chicks$Chick)
```



For most of the chicks, there are 12 repeated measures. However, there are some missing measures. For example, chick 15 only has 8 measures, chick 18 only has 2 measures. 





#3. outliers
```{r}
hist(chicks$weight)
```

```{r}
boxplot(chicks$weight)
```

```{r}
summary(chicks$weight)
```

Seems like there are some outliers present in this data 


#4. weight change overtime 

## Overall
```{r}
ggplot(data = chicks, aes(x = Time, y = weight)) + 
  geom_point() + 
  geom_smooth()
```
Overall, weight increases as time increases

```{r}

ggplot(data = chicks, aes(x = Time, y = weight, color = Diet)) + 
  geom_point() + 
  geom_smooth()
```
##individual

```{r}
gg.base <- ggplot(chicks, aes(x = Time, y = weight)) 
gg.idline <- gg.base + geom_line(aes(color = Chick, group = Chick))
gg.idline

```



#5. regression 



If we assume each row is independent

```{r}
lm1 <- lm(weight ~ Time, data = chicks)
summary(lm1)
```


```{r}
lm2 <- lm(weight ~ Diet, data = chicks)
summary(lm2)
```

```{r}
lm3 <- lm(weight ~ Time + Diet, data = chicks)
summary(lm3)
```


```{r}
anova(lm1,lm3)
```

Seems like, if we assume each row is independent, variable Time and Diet are both statistically significant with p-values less than 0.05. 











