---
 title: "EBM analysis"
 author: "Leo Zhou"
 date: "`r Sys.Date()`"
 output:
   rmdformats::downcute:
     self_contained: yes
     thumbnails: yes
     lightbox: yes
     gallery: no
     highlight: tango
   html_document:
     df_print: paged
---



<!-- # --- -->
<!-- # title: "ebm_analysis" -->
<!-- # author: "Leo Zhou" -->
<!-- # date: "7/9/2021" -->
<!-- # output: html_document -->
<!-- # --- -->



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

```{r, echo=FALSE}
print("Last run on")
format(Sys.Date(),)
```

```{r, message=FALSE}
library(ggplot2)
library(dplyr)
library(lme4)
library(MASS)
library(geepack)
library(glmmLasso)
library(car)
library(MuMIn)
library(afex)
library(blme)
library(magrittr)
library(ggeffects)
library(sjmisc)
library(splines)
```






## Functions {.tabset .tabset-fade .tabset-pills}
```{r}
lrAcc <- function(lrMod, responseVar, use.ranef=TRUE){
    ## convert response variable into a factor if it's not one
  if(!is.factor(model.frame(lrMod)[,responseVar])){
    model.frame(lrMod)[,responseVar] <- as.factor(model.frame(lrMod)[,responseVar])
  }

    ## model predictions in log-odds
  if(use.ranef){
    preds = predict(lrMod, newdata=model.frame(lrMod))
  } else{
    ## predict for average participant/item/other
        preds = predict(lrMod, newdata=model.frame(lrMod), re.form=NA)
  }
    ## transform to 0/1 prediction
    preds <- ((sign(preds)/2)+0.5)
    
    ## response variable values, transformed to 0/1
    respVarValues <- model.frame(lrMod)[,responseVar]
    ## if the response is already 0/1, just use it
    if(is.numeric(respVarValues)){
        y <- respVarValues
    } else{
        ## otherwise, transform to 0/1
        y <- (as.numeric(model.frame(lrMod)[,responseVar])-1)
    }
    
    ## how often is prediction the same as the actual response
    acc <- sum(preds==y)/length(preds)
    
    return(acc)
    
}

baselineAcc <- function(lrMod, responseVar){
    response <- model.frame(lrMod)[,responseVar]
    tab <- table(response)
    return(max(tab)/sum(tab))
}
```

## Overuse EBM {.tabset .tabset-fade .tabset-pills}

### Data cleaning
```{r}
over.df <- read.csv("C:/Users/leo.zhou/Desktop/Equity OKR/R/data/overuse_df.csv",
                    na.strings=c("","NA"))

```

```{r echo = T, results = 'hide'}
summary(over.df)
```


```{r}
hist(over.df$numerator)
```

```{r}
#there are some numerators = 2, and the denominator = 2, for regression purpose, change those to 1
#over.df[over.df$numerator==2,]
```

```{r}
#for regression purpose, set the denominator == 2 to 1
over.df$numerator[over.df$numerator == 2 ] <- 1
```

```{r}
hist(over.df$numerator)
```

```{r}
#explore numerator > 1
#over.df[over.df$numerator > 1,]
```


```{r}
#explore numerator > 1
#over.df[over.df$numerator > 0 & over.df$numerator < 1 ,]
```


```{r}
over.df$ci_measure_cat <- as.factor(over.df$ci_measure_cat)
levels(over.df$ci_measure_cat)
```



```{r echo = T, results = 'hide'}
#for regression purpose, remove all the NYU measures
over.df <- over.df[over.df$ci_measure_cat != "NYU",]
summary(over.df)
```

```{r}
#change the numerator to factor 0 and 1
over.df$numerator <- as.factor(over.df$numerator)
levels(over.df$numerator)
summary(over.df$numerator)
```


```{r}
#change the ndi to numeric
over.df$adi_natrank <- as.numeric(over.df$adi_natrank)
summary(over.df$adi_natrank)
```

```{r}
#change the ndi to numeric
over.df$adi_staternk <- as.numeric(over.df$adi_staternk)
summary(over.df$adi_staternk)
```

```{r}
#change the predicted_race to factor
over.df$predicted_race <- as.factor(over.df$predicted_race)
summary(over.df$predicted_race)
```

```{r}
#change the gender to factor
over.df$gender <- as.factor(over.df$gender)
summary(over.df$gender)
```

```{r}
over.df$over.underuse <- as.factor(over.df$over.underuse )
summary(over.df$over.underuse)
```

```{r}
over.df$ebm_measure_label <- as.factor(over.df$ebm_measure_label)
summary(over.df$ebm_measure_label)
```


```{r}
#relevel: change the reference level for predicted_race to white
over.df <- within(over.df, predicted_race <- relevel(predicted_race, ref = 6))
```




```{r}
#rescale variable age and ndi - scale at center
#if scale=TRUE, x - mean(x) / sd(x)
over.df[,"age"] <- scale(over.df[,"age"])
over.df[,"adi_natrank"] <- scale(over.df[,"adi_natrank"])
```


```{r}
hist(over.df$age)
```


```{r}
hist(over.df$adi_natrank)
```


### EDA 
Here are the variables we are interested: 
predicted_race
gender
age 
adi_natrank

#### preicted_race
```{r}
ggplot(data = over.df) +
  geom_bar(mapping = aes(x = predicted_race))
```

```{r}
summary(over.df$predicted_race)
```


```{r}
ggplot(data = over.df) +
  geom_bar(mapping = aes(x = predicted_race, fill = numerator))
```

```{r}
prop.table(table(over.df$numerator, over.df$predicted_race),1)
```


```{r}
prop.table(table(over.df$predicted_race,over.df$numerator),1)
```


#### gender

```{r}
summary(over.df$gender)
```

```{r}
ggplot(data = over.df) +
  geom_bar(mapping = aes(x = gender, fill = numerator))
```


```{r}
prop.table(table(over.df$numerator, over.df$gender),1)
```

#### age

```{r}
ggplot(data=over.df, aes(x = age)) + 
  geom_histogram()
```

```{r}
# number of NAs 
sum(is.na(over.df$age))
```



```{r}
ggplot(data=over.df, aes(x = numerator, y = age)) + 
  geom_boxplot()
```

```{r}
ggplot(data=over.df, aes(x = numerator, y = age)) + 
  geom_violin()
```

```{r}
age_summary <- over.df %>%
  group_by(numerator) %>%
  summarize(n = n(),
         mean = mean(age,na.rm=TRUE),
         sd = sd(age,na.rm=TRUE) 
         )
age_summary
```

```{r}
ggplot(data = over.df, aes(x =age, y = as.numeric(numerator) )) + 
  geom_point() +
  geom_jitter()
```



#### ndi

```{r}
ggplot(data=over.df, aes(x = adi_natrank)) + 
  geom_histogram()
```

```{r}
ggplot(data=over.df, aes(x = numerator, y = adi_natrank)) + 
  geom_boxplot()
```


```{r}
ggplot(data=over.df, aes(x = numerator, y = adi_natrank)) + 
  geom_violin()
```


```{r}
ndi_summary <- over.df %>%
  group_by(numerator) %>%
  summarize(n = n(),
         mean = mean(adi_natrank,na.rm=TRUE),
         median = median(adi_natrank,na.rm=TRUE),
         sd = sd(adi_natrank,na.rm=TRUE) 
         )
ndi_summary
```

```{r}
ggplot(data = over.df, aes(x =adi_natrank, y =numerator)) + 
  geom_point() + 
  geom_jitter()
```



#### race vs. ndi vs numerator

```{r}
ggplot(over.df, aes(x = predicted_race, y = adi_natrank)) + 
  geom_boxplot()

```

```{r}
ggplot(over.df, aes(x = predicted_race, y = adi_natrank, color = numerator)) + 
  geom_boxplot()

```

```{r}
ggplot(over.df, aes(x = age, y = adi_natrank, color = predicted_race)) + 
  geom_point() + 
  geom_smooth(method = 'lm')

```





### regression
Code and resources from Julian Faraway (2016): Extending the Linear Model with R

note from the data, each observation (row) might not be independent to each other. It is possible that one member has different meausres. Thus we should consider mised effect model with lmer(), actually glmer() that can account for the repeated measures -- all the obs are not independent. 

try glm:
```{r}
#note this model assumes all obs are independent
m0.glm <- glm(numerator ~ predicted_race + gender + age + adi_natrank,
                family = "binomial",
                data = over.df)
summary(m0.glm)

```

#### method 1 - PQL (penalized quasi likelihood)
```{r}
m0.pql <- glmmPQL(numerator ~ predicted_race + gender + age + adi_natrank,  
                random=list(member_number = ~1, ebm_measure_label = ~1),
                family = "binomial",
                data = over.df)
summary(m0.pql)

```
According to Faraway(2016), this method is based on the linearized model and rather dubious assumptions explained in 10.2, and the binary response can lead to biased estimates of the coefficients. 


#### method 2 - glmer
```{r}
#overall model
m0.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(m0.glmer)
```

```{r}
#this model does not include ndi. 
m0.glmer.race <- glmer(numerator ~  predicted_race + 
                         gender + 
                         age + 
                         (1|member_number) + 
                         (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(m0.glmer.race)
```

```{r}
#model exlude race
m0.glmer.ndi <- glmer(numerator ~ 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(m0.glmer.ndi)
```

```{r}
Anova(m0.glmer,type="II")
```

```{r}
r.squaredGLMM(m0.glmer)
```

```{r}
#diagnostic plots
#plot(m0.glmer)
```

```{r}
me1 <- ggpredict(m0.glmer, "predicted_race")
plot(me1)
```
```{r}
me2 <- ggpredict(m0.glmer, "adi_natrank")
plot(me2)
```

```{r}
exp(coef(summary(m0.glmer))[ , "Estimate"])
```

#### method 3 - Bayesian MEMs
```{r}
m0.bayes <- bglmer(numerator ~  predicted_race + 
                     gender + 
                     age + 
                     adi_natrank + 
                     (1|member_number) + 
                     (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df,
                control=glmerControl(optimizer = "bobyqa"))
summary(m0.bayes)
```




### Test for significance of variables



```{r}
#lrMod1.pb <- mixed(numerator ~ predicted_race + gender + age + adi_natrank + (1|member_number) + (1|ebm_measure_label),
#                data=over.df, 
#                family="binomial", method='LRT', args_test=list(nsim=1000), 
#                control=glmerControl(optimizer = "bobyqa"))

#summary(lrMod1.pb)
```
#### test random effect
```{r}
m0.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age +
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df)

m0.glmer.r <- glmer(numerator ~  predicted_race + 
                      gender + 
                      age + 
                      adi_natrank + 
                      (1|member_number),
                family = binomial(link="logit"),
                data = over.df)

anova(m0.glmer, m0.glmer.r)
```
Thus, we should include the random effect term: (1|ebm_measure_label)


control = glmerControl(tolPwrss=1e-3)
control=glmerControl(optimizer = "bobyqa")
#### test interactions
```{r}
# test the effect predicted_race:ndi
m0.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df,
                control = glmerControl(tolPwrss=1e-3)
                )

m0.glmer.racendi <- glmer(numerator ~  predicted_race + 
                            gender + 
                            age + 
                            adi_natrank + 
                            predicted_race:adi_natrank + 
                            (1|member_number) + 
                            (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df,
                control = glmerControl(tolPwrss=1e-3))


anova(m0.glmer, m0.glmer.racendi)
```
 
```{r}
summary(m0.glmer.racendi)
```

```{r}
Anova(m0.glmer.racendi, type =  "II")
```

No need for the interaction: predicted_race:adi_natrank


```{r}
# test the effect gender:ndi
m0.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age +
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df)

m0.glmer.genderndi <- glmer(numerator ~  predicted_race + 
                              gender + 
                              age +
                              adi_natrank + 
                              gender:adi_natrank + 
                              (1|member_number) + 
                              (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df)


anova(m0.glmer, m0.glmer.genderndi)
```
No need for the interaction: gender:adi_natrank




```{r}
# test the effect age:ndi
m0.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age +  
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df)

m0.glmer.agendi <- glmer(numerator ~  predicted_race + 
                           gender + 
                           age +
                           adi_natrank + 
                           age:adi_natrank + 
                           (1|member_number) + 
                           (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = over.df)


anova(m0.glmer, m0.glmer.agendi)
```
the interaction age:ndi appears to be signicant, take a look at the summary output

```{r}
summary(m0.glmer.agendi)
```


Visualize age:ndi interaction

```{r}
ggplot(over.df, aes(x = age, y = adi_natrank, color = numerator)) + 
  geom_point() + 
  geom_smooth(method='lm',size=1.5)
```




### goodness-of-fit

Test correlation between ndi and race
```{r}
model.lm <- lm(adi_natrank ~ predicted_race, data = over.df)
summary(model.lm)

```


```{r}
rsq <- summary(model.lm)$r.squared
rsq

```

```{r}
kruskal.test(adi_natrank ~ predicted_race, data = over.df)
```

```{r}
r.squaredGLMM(m0.glmer)
```

```{r}
summary(m0.glmer)$varcor
```

```{r}
# entire model: fixed + random
lrAcc(m0.glmer, 'numerator')
```


```{r}

#baseline is the intercept only model and random effets
baselineAcc(m0.glmer, 'numerator')
```
```{r}
#fixed effects only
lrAcc(m0.glmer, 'numerator', use.ranef=FALSE)
```


## Underuse EBM {.tabset .tabset-fade .tabset-pills}
### Data cleaning
```{r}
under.df <- read.csv("C:/Users/leo.zhou/Desktop/Equity OKR/R/data/underuse_df.csv",
                    na.strings=c("","NA"))
nrow(under.df)
```

```{r}
hist(under.df$numerator)
```

```{r}
under.df$ci_measure_cat <- as.factor(under.df$ci_measure_cat)
levels(under.df$ci_measure_cat)
```




```{r}
#change the numerator to factor 0 and 1
under.df$numerator <- as.factor(under.df$numerator)
levels(under.df$numerator)
summary(under.df$numerator)
```


```{r}
under.df$denominator <- as.factor(under.df$denominator )
summary(under.df$denominator)
```

```{r}
#change the ndi to numeric
under.df$adi_natrank <- as.numeric(under.df$adi_natrank)
summary(under.df$adi_natrank)
```


```{r}
#change the predicted_race to factor
under.df$predicted_race <- as.factor(under.df$predicted_race)
summary(under.df$predicted_race)
```

```{r}
#change the gender to factor
under.df$gender <- as.factor(under.df$gender)
summary(under.df$gender)
```

```{r}
under.df$over.underuse <- as.factor(under.df$over.underuse )
summary(under.df$over.underuse)
```

```{r}
#relevel: change the reference level for predicted_race to white
under.df <- within(under.df, predicted_race <- relevel(predicted_race, ref = 6))
```



```{r}
#rescale variable age and ndi - scale at center
#if scale=TRUE, x - mean(x) / sd(x)
under.df[,"age"] <- scale(under.df[,"age"])
under.df[,"adi_natrank"] <- scale(under.df[,"adi_natrank"])
```


```{r}
hist(under.df$age)
hist(under.df$adi_natrank)
```






### EDA 
#### predicted_race 
```{r}
ggplot(data = under.df) +
  geom_bar(mapping = aes(x = predicted_race))
```

```{r}
summary(under.df$predicted_race)
```


```{r}
ggplot(data = under.df) +
  geom_bar(mapping = aes(x = predicted_race, fill = numerator))
```

```{r}
prop.table(table(under.df$numerator, under.df$predicted_race),1)
```


```{r}
prop.table(table(under.df$predicted_race,under.df$numerator),1)
```




#### gender 
```{r}
summary(under.df$gender)
```

```{r}
ggplot(data = under.df) +
  geom_bar(mapping = aes(x = gender, fill = numerator))
```


```{r}
prop.table(table(under.df$numerator, under.df$gender),1)
```



#### age 
```{r}
ggplot(data=under.df, aes(x = age)) + 
  geom_histogram()
```

```{r}
# number of NAs 
sum(is.na(under.df$age))
```



```{r}
ggplot(data=under.df, aes(x = numerator, y = age)) + 
  geom_boxplot()
```

```{r}
ggplot(data=under.df, aes(x = numerator, y = age)) + 
  geom_violin()
```

```{r}
age_summary <- under.df %>%
  group_by(numerator) %>%
  summarize(n = n(),
         mean = mean(age,na.rm=TRUE),
         median = median(age,na.rm=TRUE),
         sd = sd(age,na.rm=TRUE) 
         )
age_summary
```

```{r}
ggplot(data = under.df, aes(x =age, y = as.numeric(numerator) )) + 
  geom_point() + 
  geom_jitter()
```

#### ndi 
```{r}
ggplot(data=under.df, aes(x = adi_natrank)) + 
  geom_histogram()
```

```{r}
ggplot(data=under.df, aes(x = numerator, y = adi_natrank)) + 
  geom_boxplot()
```


```{r}
ggplot(data=under.df, aes(x = numerator, y = adi_natrank)) + 
  geom_violin()
```


```{r}
ndi_summary <- under.df %>%
  group_by(numerator) %>%
  summarize(n = n(),
         mean = mean(adi_natrank,na.rm=TRUE),
         median = median(adi_natrank,na.rm=TRUE),
         sd = sd(adi_natrank,na.rm=TRUE) 
         )
ndi_summary
```

```{r}
ggplot(data = under.df, aes(x =adi_natrank, y =numerator)) + 
  geom_point() + 
  geom_jitter()
```

### regression

#### method 1 - PQL (penalized quasi likelihood)
```{r}
mU.pql <- glmmPQL(numerator ~ predicted_race + 
                    gender + 
                    age + 
                    adi_natrank,  
                random=list(member_number = ~1, ebm_measure_label = ~1),
                family = "binomial",
                data = under.df)
summary(mU.pql)

```



#### method 2 - glmer
```{r}
mU.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df,
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(mU.glmer)
```
```{r}
Anova(mU.glmer,type="II")
```


```{r}
r.squaredGLMM(mU.glmer)
```


```{r}
me2 <- ggpredict(mU.glmer, "predicted_race")
plot(me2)
```

```{r}
me2 <- ggpredict(mU.glmer, "adi_natrank")
plot(me2)
```


```{r}
exp(coef(summary(mU.glmer))[ , "Estimate"])
```




#### method 3 - Bayesian MEMs
```{r}
mU.bayes <- bglmer(numerator ~  predicted_race + 
                     gender + 
                     age + 
                     adi_natrank + 
                     (1|member_number) + 
                     (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df,
                control=glmerControl(optimizer = "bobyqa"))
summary(mU.bayes)
```


### Test for significance of variables

#### test random effect
```{r}
mU.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df)

mU.glmer.r <- glmer(numerator ~  predicted_race + 
                      gender + 
                      age + 
                      adi_natrank + 
                      (1|member_number),
                family = binomial(link="logit"),
                data = under.df)

anova(mU.glmer, mU.glmer.r)
```
Thus, we should include the random effect term: (1|ebm_measure_label)


#### test interactions
```{r}
# test the effect predicted_race:ndi
mU.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df)

mU.glmer.racendi <- glmer(numerator ~  predicted_race + 
                            gender + 
                            age + 
                            adi_natrank + 
                            predicted_race:adi_natrank + 
                            (1|member_number) + 
                            (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df,
                control = glmerControl(tolPwrss=1e-3))


anova(mU.glmer, mU.glmer.racendi)
```
No need for the interaction: predicted_race:adi_natrank


```{r}
# test the effect gender:ndi
mU.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df)

mU.glmer.genderndi <- glmer(numerator ~  predicted_race + 
                              gender + 
                              age +
                              adi_natrank + 
                              gender:adi_natrank + 
                              (1|member_number) + 
                              (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df)


anova(mU.glmer, mU.glmer.genderndi)
```
No need for the interaction: gender:adi_natrank for the underuse model 




```{r}
# test the effect age:ndi
mU.glmer <- glmer(numerator ~  predicted_race + 
                    gender + 
                    age + 
                    adi_natrank + 
                    (1|member_number) + 
                    (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df)

mU.glmer.agendi <- glmer(numerator ~  predicted_race + 
                           gender + 
                           age +
                           adi_natrank + 
                           age:adi_natrank + 
                           (1|member_number) + 
                           (1|ebm_measure_label),
                family = binomial(link="logit"),
                data = under.df)


anova(mU.glmer, mU.glmer.agendi)
```
for the underuse model, interaction age:ndi does not appear to be significant. 



### goodness-of-fit


Test correlation between ndi and race
```{r}
model.lm.U <- lm(adi_natrank ~ predicted_race, data = under.df)
summary(model.lm.U)

```


```{r}
rsqU <- summary(model.lm.U)$r.squared
rsqU

```

```{r}
kruskal.test(adi_natrank ~ predicted_race, data = under.df)
```


```{r}
r.squaredGLMM(mU.glmer)
```
Marginal R2 provides
the variance explained only by fixed effects and conditional R2 provides the variance
explained by the entire model, i.e., both fixed effects and random effects
```{r}
summary(mU.glmer)$varcor
```

```{r}
#fixed + random
lrAcc(mU.glmer, 'numerator')
```


```{r}
#baseline is the intercept only model: intercept + random
baselineAcc(mU.glmer, 'numerator')
```

```{r}
#fix effects only
lrAcc(m0.glmer, 'numerator', use.ranef=FALSE)
```


```{r}
r.squaredGLMM(mU.bayes)

```

```{r}
lrAcc(mU.bayes, 'numerator')
baselineAcc(mU.bayes, 'numerator')
lrAcc(m0.bayes, 'numerator', use.ranef=FALSE)
```









## Resources and citations: {.tabset .tabset-fade .tabset-pills}
Julian Faraway (2016): Extending the Linear Models with R

https://www.ssc.wisc.edu/sscc/pubs/MM/MM_DiagInfer.html


http://people.linguistics.mcgill.ca/~morgan/book/mixed-effects-logistic-regression.html#ref-Chung2013likelihood




Marginal R2 provides the variance explained only by fixed effects
Conditional R2 provides the variance explained by the entire model, i.e., both fixed effects and random effects. Citation: 
Nakagawa, S., Schielzeth, H. (2013) A general and simple method for obtaining R<U+00B2> from Generalized Linear Mixed-effects Models. Methods in Ecology and Evolution 4: 133<U+2013>142


